---
title: "hw03-Sizhuo(Cindy)-Liu"
author: "Sizhuo (Cindy) Liu"
date: "March 17, 2018"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, fig.path = 'images/')
```

###2. Pipelines and Redirection
####Piping

#####2.1 Write a pipeline to obtain the unique team names, and redirect the output to a text file team-names.txt
```{bash}
# select second column (team) 
# sort by team 
# list unique occurrences based on team name 
# redirect the output to a new textfile 
cut -d ',' -f 2 nba2017-roster.csv | sort | sort -u > team-names.txt

# display the first five lines of the created file
head -5 team-names.txt
```

#####2.2 Write a pipeline to obtain the unique positions, and redirect the output to a text file position-names.txt
```{bash}
# select third column (position) 
# sort by position
# list unique occurrences based on position name 
# redirect the output to a new textfile 
cut -d ',' -f 3 nba2017-roster.csv | sort | sort -u > position-names.txt

# display the first five lines of the created file
head -5 position-names.txt
```

#####2.3 Write a pipeline to obtain the counts (i.e. frequencies) of the different experience values, displayed from largest to smallest (i.e. descending order). Redirect the output to a text file experience-counts.txt
```{bash}
# select seventh column (experience) 
# list unique occurrences based on experience value 
# get rid of the first line (column name)
# display counts of unique values in descending order
# redirect the output to a new textfile 
cut -d ',' -f 7 nba2017-roster.csv | sort -n | tail +2 | uniq -c > experience-counts.txt

# display the first five lines of the created file
head -5 experience-counts.txt
```

#####2.4 Use output redirection commands to create a CSV file LAC.csv containing data for the LAC team (Los Angeles Clippers), including column names.
```{bash}
# filter out data for LAC players with column names
# redirecct the output to a new csv file
grep -E 'LAC|player' nba2017-roster.csv > LAC.csv 

# display the content of the created file
cat LAC.csv
```

#####2.5 Write a pipeline to display the age frequencies of LAL players. The first column corresponds to count, the second column corresponds to age
```{bash}
# filter out all rows representing LAL players 
# select sixth column (age) 
# list unique occurrences based on age
# get rid of the first line (column name)
# display counts of unique values in descending order
grep -E 'LAL|player' nba2017-roster.csv | cut -d ',' -f 6 | sort -n | tail +2 | uniq -c 

```

#####2.6 Write a pipeline to find the number of players in CLE (Cleveland) team; the output should be just the number of players
```{bash}
# filter out all rows representing CLE players 
# count the number of lines
grep --word-regexp 'CLE' nba2017-roster.csv | wc -l
```

#####2.7 Write pipelines to create a CSV file gsw-height-weight.csv that contains the player, height and weight of GSW players and includes column names
```{bash}
# filter out all rows representing GSW players
# select columns corresponding to player names, heights, and weights
# redirect the output to a new csv file
grep -E 'GSW|player' nba2017-roster.csv | cut -d ',' -f 1,4,5 > gsw-height-weight.csv

# display the content of the new csv file
cat gsw-height-weight.csv
```

#####2.8 Write pipelines to create a file top10-salaries.csv containing the top10 player salaries, arranged by salary from largest to smallest. Your CSV file should include column names
```{bash}
cut -d ',' -f 1,8 nba2017-roster.csv | sort -r -k2 -n -t, | head -10 > top10-salaries.csv
sed -i '' '1i\ 
"player","salary"
' top10-salaries.csv
cat top10-salaries.csv
```
```{bash}
grep 'player' nba2017-roster.csv | cut -d ',' -f 1,8 > top10-salaries.csv
cut -d ',' -f 1,8 nba2017-roster.csv | sort -r -k2 -n -t, | head -10 >> top10-salaries.csv
cat top10-salaries.csv
```

####Source Binomial Functions
```{r}
source(file='./code/binomial-functions.R')
```


#####Assume that the “successful” event is getting a “six” when rolling a die. Consider rolling a fair die 10 times. Use bin_probability() to find the probability of getting exactly 3 sixes
```{r}
bin_probability(trials=10, success=3, prob=1/6)
```

#####Use bin_distribution() to obtain the distribution of the number of “sixes” when rolling a loaded die 10 times, in which the number “six” has probability of 0.25. Make a plot of this distribution
```{r}
#ask
bin_distribution(trials=10, prob=0.25)
library(ggplot2)
ggplot(bin_distribution(trials=10, prob=0.25), aes(x=success,y=probability)) +
  geom_point() +
  ggtitle('Probability Distribution of the Number of "sixes" When Rolling a Die 10 Times')
```

#####Use bin_probability(), and a for loop, to obtain the probability of getting more than 3 heads in 5 tosses with a biased coin of 35% chance of heads
```{r}
bin_prob <- rep(0, 2)
k <- 3
for (i in 1:2) {
  k <- k + 1
  bin_prob[i] <- bin_probability(trials=5, success=k, prob=0.35)
  
}
sum(bin_prob)
```

#####Use bin_distribution() to obtain the probability distribution of the number of heads when tossing a loaded coin 15 times, with 35% chance of heads. Make a plot of this distribution
```{r}
bin_distribution(trials=15, prob=0.35)
ggplot(bin_distribution(trials=15, prob=0.35), aes(x=success,y=probability)) +
  geom_point() +
  ggtitle('Probability Distribution of the Number of Heads When Tossing a Coin 15 Times')
```









