---
title: "Warmup 2"
author: Sizhuo(Cindy) Liu
date: February 6,2018
output: github_document
---
###Getting the Data File
```{r}
load("/Users/sizhuoliu/nba2017-salary-points.RData")
```
###Data Processing

* converting unit of salary
```{r}
salary_millions <- salary/1000000
round(salary_millions,digits=2)
```

* Renaming factors
```{r}
position_fac<-factor(position,
                     levels=c("C","SF","PF","SG","PG"),
                     labels=c("center","small_fwd","power_fwd","shoot_guard","power_guard"))
position_fac
```
####Quantitative Variable

* Descriptive Statistics of Salary
```{r}
summary(salary)
```
```{r}
#average:typical value
mean(salary)
#standard deviation:spread
sd(salary)
#minimum value
min(salary)
#maximum value 
max(salary)
#median value
median(salary)
#quartiles
quantile(salary,c(0.25,0.5,0.75))
```

* Descriptive Statistics of Points
```{r}
#average:typical value
mean(points)
#standard deviation: spread
sd(points)
```

* Graphs
```{r}
#Distribution
hist(salary)
hist(points)
boxplot(salary)
boxplot(points)
```
```{r}
#kernel density plot
plot(density(salary))
plot(density(points))
```

**Description of Overall Patterns**

* For both the plots for salary and for points, the density is the highest at the lower end, and gradually fade away as the salary/score increases. 
* There are no striking deviations in either plot.

####Qualitative Variable

* Creating Factors
```{r}
#turn team into a R factor
team_fac<-factor(team)
team_fac
```
```{r}
table(team_fac)
frequency<-table(team_fac)
relative_frequency<-frequency/441
relative_frequency
```
###Scatterplot of Points and Salary

* Scatterplot Color-coded Based on Team
```{r}
plot(points,salary_millions,col=team_fac)
```

* Scatterplot Color-coded Based on Position
```{r}
plot(points,salary_millions,col=position_fac)
```

###Correlation Between Points and Salary

* Mean of points/Points
```{r} 
points_mean <- (1/441)*sum(points)
points_mean
```

* Mean of Salary
```{r}
salary_mean <- (1/441)*sum(salary)
salary_mean
```

* Variance of points/Points
```{r}
n <- length(player)
points_var <- (1/(n-1))*sum((points-points_mean)^2)
points_var
```

* Variance of Salary
```{r}
salary_var <- (1/(n-1))*sum((salary-salary_mean)^2)
salary_var
```

* Standard Deviation of points/Points
```{r}
points_sd <-sqrt(points_var)
points_sd
```

* Standard Deviation of Salary
```{r}
salary_sd <- sqrt(salary_var)
salary_sd
```

* Covariance of points/Points and Salary
```{r}
points_salary_cov <- (1/(n-1))*sum((points-points_mean)*(salary-salary_mean))
points_salary_cov
```

* Correlation Between points and Salary
```{r}
points_salary_corr <- points_salary_cov / ((points_sd)*(salary_sd))
points_salary_corr
```
###Comments and Reflections

* The one part that is relatively hard for me is plotting the graphs. Although we have already done that in labs, I found them difficult because I think that it is tricky to set up the graph in a way so that they are organized and easily understandable. 

* The part that is relatively easy for me was calculating the mean, variance, standard deviation, etc. of the data. Since I know the formulas for calculating these things, it is very straightforward to apply the calculations in R. 

* I needed help with this assignment as I was stuck on some parts such as using the sigma notation correctly. The issue was solved when I attended my GSI's office hour, where she offered hints to help me solve the problem. 

* It took in total around 3 hours (mainly because I spent a very long time on one single part). 

* The most time-consuming part was calculating the mean/sd/variance, etc. as I had to double check to make sure I got the right answers. 

* The one thing that I think that I still haven't fully grasped is the idea of factor vs. general vectors. Although I know how to apply them, I still do not fully understand the reasoning behind using one as opposed to the other. 

* The most frustrating part was figuring out how to use the sigma notation as I spent the most time on this and had to look for information online. 


