---
title: "HW02 - Shot Charts"
author: "Sizhuo (Cindy) Liu"
date: "February 28, 2018"
output: github_document
---                 
###5. Summary Tables
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/')
```

```{r}
library(dplyr)
library(ggplot2)
```

####5.1 Total Shots by Players
```{r}
# total number of shots made by players, arranged in descending order
shots_data <- read.csv('../data/shots-data.csv', stringsAsFactors = FALSE)

shots_data %>%
  group_by(name) %>%
  summarise(total_shots = length(shot_type)) %>%
  arrange(desc(total_shots))
```

####5.2 Effective Shooting Percentage
```{r}
# effective shooting percentage by player
merge(shots_data %>%
  group_by(name) %>%
  summarise(total_shots = length(shot_type)),
  shots_data%>%
  group_by(name) %>%
  filter(shot_made_flag == 'made shot') %>%
  summarise(made_shots = length(shot_made_flag))) %>%
  mutate(perc_made = made_shots / total_shots) %>%
  arrange(desc(perc_made))
```

```{r}
# 2PT effective shooting percentage by player
merge(shots_data %>%
  group_by(name) %>%
    filter(shot_type == '2PT Field Goal') %>%
    summarise(total_shots = length(shot_type)),
  shots_data%>%
  group_by(name) %>%
  filter(shot_made_flag == 'made shot') %>%
    filter(shot_type == '2PT Field Goal') %>%
    summarise(made_shots = length(shot_type))) %>%
  mutate(perc_made = made_shots / total_shots) %>%
  arrange(desc(perc_made))
```

```{r}
# 3PT effective shooting percentage by player
merge(shots_data %>%
  group_by(name) %>%
    filter(shot_type == '3PT Field Goal') %>%
    summarise(total_shots = length(shot_type)),
  shots_data%>%
  group_by(name) %>%
  filter(shot_made_flag == 'made shot') %>%
    filter(shot_type == '3PT Field Goal') %>%
    summarise(made_shots = length(shot_type))) %>%
  mutate(perc_made = made_shots / total_shots) %>%
  arrange(desc(perc_made))
```

####6. Shooting Distance
#####6.1 dplyr Table
```{r}
# create new tibble using dplyr functions
tibble1 <- merge(shots_data %>%
  group_by(shot_distance) %>%
  summarise(total_shot1 = length(shot_made_flag)),
  shots_data %>%
  group_by(shot_distance) %>%
    summarise(made_shot1 = sum(shot_made_flag == 'made shot'))) %>%
      group_by(shot_distance) %>%
  summarise(made_shot_prop = made_shot1 / total_shot1) 
tibble1
```

#####6.2 ggplot
```{r scatterplot}
# create a scatterplot based on the tibble in 6.1
ggplot(data = tibble1, aes(x=shot_distance, y=made_shot_prop)) +
  geom_point()
```

#####Observation of the Pattern of the Graph
* The scatterplot shows that shots made at the distances between 0 and 2 away from the goal have the highest chance of being successful, with a proportion of around 50% to 80%. The percentage drops below 50% when the distance goes above 3 to around 25%-50%. The chance that a shot is being made successful at a distance above 33 drops drastically to almost zero. 
* We can confirm that the shorter the distance the more effective the shots as the made_shot_proportion becomes smaller and smaller as the shot distance increases. 
* According to the graph, the estimated distance threshold beyond which the chance of making a successful shot is null is anything above 30. 
* Distances of 0-2, 12, 16, 18, and 21 tend to have a made_shot_prop of 50% or more. 

####7. Total Number of Shots by Minute of Occurrence
```{r}
# create a facetted graph of the total number of shots by minute of occurrence

# reorganize data
shots1 <-shots_data %>%
  select(minute,name) %>%
  filter(name== 'Andre Iguodala') %>%
  group_by(minute) %>%
  summarise (number_of_shots = length(name == name)) %>%
  mutate(name = 'Andre Iguodala')
shots2 <-shots_data %>%
  select(minute,name) %>%
  filter(name== 'Draymond Green') %>%
  group_by(minute) %>%
  summarise (number_of_shots = length(name == name)) %>%
  mutate(name = 'Draymond Green')
shots3 <-shots_data %>%
  select(minute,name) %>%
  filter(name== 'Kevin Durant') %>%
  group_by(minute) %>%
  summarise (number_of_shots = length(name == name)) %>%
  mutate(name = 'Kevin Durant')
shots4 <-shots_data %>%
  select(minute,name) %>%
  filter(name== 'Klay Thompson') %>%
  group_by(minute) %>%
  summarise (number_of_shots = length(name == name)) %>%
  mutate(name = 'Klay Thompson')
shots5 <-shots_data %>%
  select(minute,name) %>%
  filter(name== 'Stephen Curry') %>%
  group_by(minute) %>%
  summarise (number_of_shots = length(name == name)) %>%
  mutate(name = 'Stephen Curry')
shots_overall <- rbind(shots1,shots2,shots3,shots4,shots5)
shots_overall
```


```{r facetted-graph}
# plot facetted graph
ggplot(data = shots_overall,aes(x = minute, y =number_of_shots)) + 
  geom_rect(xmin = 0,xmax = 12,ymin = 0,ymax = 60, alpha = 0.005, col = 'grey96') +
  geom_rect(xmin = 24,xmax = 36,ymin = 0,ymax = 60, alpha = 0.005, col = 'grey96') +
  geom_point(col = 'dodgerblue2', alpha = 0.8) +
  geom_path(col = 'dodgerblue2', alpha = 0.5) +
  scale_x_continuous()+
  facet_wrap(~name) +
  theme_minimal()
```

















